---
title: "How to Sell Out for Popularity in Music"
subtitle: "Data Driven Recommendations for Making Songs to Limit Spotify Skips"
author: <a href=mailto:kmhurchla@gmail.com>kmhurchla@gmail.com</a>
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook:
    code_folding: hide
    fig_height: 4
    fig_width: 6
    toc: yes
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

<style>
@import url('https://fonts.googleapis.com/css2?family=Roboto:wght@300&display=swap');
</style>

```{css, echo=FALSE}
* {
  font-family: 'Roboto', sans-serif;
}

h1, .h1  {
    margin-top: 70px;
}
```
<img src=`r "../img/headphone.png"` width="200"> <img src=`r "../img/fast.png"` width="200">

Navigate easily by clicking the blue button in the upper left corner for shortcuts around this EDA notebook.

```{r setup, include = FALSE, echo = FALSE, message = FALSE, cache = FALSE, fig.show='asis'}
# Disable scientific notation in double numbers, because instrumentalness otherwise did not test plotting accurately (plotted at zero)
options(scipen = 999)

knitr::opts_chunk$set(
  echo = F,
  warning = F,
  message = F,
  include = T,
  cache = F,
  fig.width = 6, 
  fig.height = 4,
  out.width = "100%", 
  out.height = "100%"
  )

knitr::clean_cache(
  clean = FALSE, 
  path = knitr::opts_chunk$get("cache.path")
  )
```

```{r install Baltimore City ggiteam package}
# We'll skip this and use other color options because I continue to get errors even with attempting to install it remotely; possibly a permissions related error?
# package object 'ggiteam' is a color standard styling pckg for City of Balt. (see https://rdrr.io/github/city-of-baltimore/ggiteam/src/R/ggiteam.R)
#install.packages("remotes")
#remotes::install_github("city-of-baltimore/ggiteam")
```

```{r load_packages}
library(tidyverse)
library(dplyr) #loaded this explicitly in addition to tidyverse on recommendation that sometimes pipe won't work with tidyversy loaded alone, although package is included with tidyverse.
library(lubridate)
library(sf)
library(leaflet)
#library(ggiteam)
#library(package)
#library(kableExtra)
library(htmltools)
library(DT)
# Note quotes used in install of package below, becuase it otherwise would error and cannot be found and in RStudio packages list
#install.packages("fmsb")
library(fmsb)
#library(viridis)
library(RColorBrewer)
library(ggplot2)
```

# Exploratory Data Analysis notebook

```{r load-data-from-csv}
track_features <- read_csv("../data/raw/track_features/tf_mini.csv")
skips <- read_csv("../data/raw/training_set/log_mini.csv")
```

```{r sort desc by unique track count}
# counts of number of times each track (by ID) appears in training dataset and sort table by counts descending

skips %>%
  count(track_id_clean) %>%
  arrange(desc(n))
```

```{r sample}
# sample data subset; add percent not skipped
track_plays_and_skips <- skips %>%
  group_by(track_id_clean) %>%
  summarise(
    plays = n(),
    skip_1 = sum(skip_1),
    skip_2 = sum(skip_2),
    skip_3 = sum(skip_3),
    not_skipped = sum(not_skipped)
  ) %>%
  # Take only the tracks played 50 or more times.
  filter(plays >= 50) %>%
  # Add percentage of times each was not skipped of its total times played.
  mutate(not_skipped_pct = not_skipped / plays)
```

```{r join-features-skips}
# Join features to plays_skips.
track_plays_skips_features <- track_plays_and_skips %>% 
  left_join(track_features, by = c("track_id_clean" = "track_id")) 
```

```{r glimpse joined plays_skips to features table}
glimpse(track_plays_skips_features)
```
Tracks that are likely to be podcasts or audio books must be removed, based on a speechiness above 0.66 according to Spotify documentation. 
(Below .33 music/little speech; .33-.66 both speech and music.)
Under "AudioFeaturesObject" view "Speechiness" at source: https://developer.spotify.com/documentation/web-api/reference/#endpoint-search

```{r filter tracks for music}
# Commented out due to error: Error in music_plays_skips_features(.) : could not find function "music_plays_skips_features"
# After analyzing data, there are none above this range which appear in our training tracks (only in actual features table), so we can debug this at another time for future refreshing of analysis.
#music_plays_skips_features <- track_plays_skips_features %>%
#  filter(speechiness <= 0.66) %>%
#  music_plays_skips_features
```

### Initial Scatterplots by Features

Let's take some peeks for relationships of playing songs through with some of our audio features.

```{r track_plays_skips_features-plot-energy, dev='png', fig.show='asis'}
p <- track_plays_skips_features %>%
  ggplot(aes(x = energy, y = not_skipped_pct)) +
  geom_point()
# Add an understandable plot title and clean axis labels
p + ggtitle("Percentage of Times each Track was Skipped of its Times Played \n by Energy of Tracks") +
  xlab("Energy") + ylab("Not Skipped %") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r track_plays_skips_features-facets-plots, plot.width = 3, out.width= "100%", dev='png', fig.show='asis'}
track_plays_skips_features %>%
  select(
    not_skipped_pct,
    energy,
    bounciness,
    beat_strength,
    danceability,
    flatness,
    instrumentalness,
    speechiness,
    acousticness,
    liveness
    
    
  ) %>%
  gather(key = metric_name, value = metric, -not_skipped_pct) %>%
  ggplot(aes(x = metric, y = not_skipped_pct)) +
  geom_point() +
  facet_wrap(~metric_name) +
# Add an understandable plot title and clean axis labels
  ggtitle("Percentage of Times each Track was Not Skipped of its Times Played \n by Various Audio Metrics") +
  xlab("metric") + ylab("Not Skipped %") +
  theme(plot.title = element_text(hjust = 0.5))
```
*Conclusion*: No strong/obvious visually apparant patterns of correlation can be seen in these scatterplots to address our inquiry of music's features with likelihood of being skipped.

**SHAPE OF TRACKS**

Next, we're going to look at most and least skipped tracks on a radial chart (aka a spider, web, polor or radar charts/graphs).
Before we do so, it's important to note this graph type may make comparisons more difficult than linear styles in some situations, and you can read more about that here: https://www.data-to-viz.com/caveat/spider.html

Our situation is already normalized where the axis of each audio metric is the same which lends itself to more easily understandable and visually accurate view, for example it will be read more logically and accurately than if your flatness scale was 0 to 1 but energy was 0 to 10, etc.

```{r plot radar chart least skipped by percentage}
# The fmsb library is useful to build a basic radar chart
# Very specific input format is required so first we need to revisit and reshape our data track_plays_skips_features
# According to https://www.r-graph-gallery.com/142-basic-radar-chart.html: "Each row must be an entity. Each column is a quantitative variable. 
# First 2 rows provide the min and the max that will be used for each variable."
# Our project goal is to show our musician client how to "sell out" i.e. what kind of music to make to achieve the most plays with least skips, so we'll start with the tail of the dataset we've isolated,
# since it already is sorted descending by skips, this will give us the last 10 rows which all were played more than 50 times but were skipped the least of the songs played that many times or more.

tracks_least_skipped_features <- tail(track_plays_skips_features)
tracks_least_skipped_features
```

## Radar Requirements Checklist & Data Preparations

How far we are from what is required for a radar chart?

**1. Each row must be an entity.**
    + Each row represents a unique track (track_id_clean).

**2. Each column is a quantitative variable.**
    + Currently, columns contain summarized aggregations of its plays/skips, 
    + and columns for other quantitative attributes like energy, dancability, etc. 
    + We need to remove or drop any columns that won't show in the graph, or are not quantitative.

**3. First 2 rows provide the min and the max** that will be used for each variable.
    + To set these range(s), we'll refer to both Spotify Developer documentation and review min/max of our data.
    + Goals are to keep a single consistent range as much as possible, and
    + provide logical and accurate ranges, which I've found has required music terminology research in some cases.


```{r check data types of variable columns}
# Let's check data types.
glimpse(tracks_least_skipped_features)
```
**2. Again, each column is a quantitative variable:** _Non-quantitative or columns unnecessary for this chart will be removed:_

*  $ not_skipped_pct 
  + _We're sorting by this descending, and there's no real use in visualizing it since our goal is to visualize the attributes of audio that makes it popular, i.e. not skipped frequently._
*  $ track_id_clean 
  + _It seems according to instructions this is no longer needed; row represent an entity but id is unnecessary for chart (since songs are anonymized it really doesn't matter here.)_
*  $ plays 
  + _(We're already representing this by descending order of the percentage, which will also order our separate charts dispayed)._
*  $ skip_1 (and 2,3) 
  + _(for our purposes, we'll simplify to only analyze skipped or not skipped, as normalized by the percentage aggregation)_
*  $ not_skipped 
  + _(redundant; same reasons as above)_
*  $ duration 
  + _let's try it first, scale is vastly different than other columns and may be confusing--we'll need to explicitly label each axis to try)_
*  $ release_year 
  + _ditto duration; I think this will be better as a caption/detail attribute, but we can try first._
*  $ us_popularity_estimate 
  + _ditto duration; this in other published models has shown to be the most indicative of success (least skips), but the scale is very large so we may have trouble with that._
*  $ mode 
  + _character type: major, minor. Move to a caption label or ignore._
*  $ acoustic_vector_0 (and 1-7) 
  + _Remove for separate plotting; we'll assign a new variable table for this analysis so we can easily return to work with these separately._
*  $ dyn_range_mean 
  + _removed as redundant and rolled up into energy; also has a very distinct 0 to 35 scale range different from other features._
  
Everything else is a double type and we'll try with labelled axis on each, and determine if it's too much, too busy, and go from there.
 
```{r subset least skipped dataframe to keep only desired column variables}
# Subset remove column (-c); do not use quotes around column names in subset.
# Commented out due to error I could not yet debug; trying a different method instead "Error in -c(not_skipped_pct, track_id_clean, plays, skip_1, skip_2, skip_3, not_skipped, : invalid argument to unary operator"
#tracks_least_skipped_radar = subset(tracks_least_skipped_features, -c(not_skipped_pct,track_id_clean,plays,skip_1,skip_2,skip_3,not_skipped,dyn_range_mean,mode,acoustic_vector_0,acoustic_vector_1,acoustic_vector_2,acoustic_vector_3,acoustic_vector_4,acoustic_vector_5,acoustic_vector_6,acoustic_vector_7))
```
```{r drop-colmns}
# Drop least skipped dataframe to keep only desired column variables.
# Note quotes are required around objects to call columns by name in this format.
drop <- c("not_skipped_pct","track_id_clean","plays","skip_1","skip_2","skip_3","not_skipped","dyn_range_mean","mode","acoustic_vector_0","acoustic_vector_1","acoustic_vector_2","acoustic_vector_3","acoustic_vector_4","acoustic_vector_5","acoustic_vector_6","acoustic_vector_7")
tracks_least_skipped_radar = tracks_least_skipped_features[,!(names(tracks_least_skipped_features) %in% drop)]
```
  
```{r glimpse updated radar dataframe}
glimpse(tracks_least_skipped_radar)
```
__...back to Radar Requirements Checklist__

**3. First 2 rows provide the min and the max** that will be used for each variable.

Let's see, we'll need to determine and assign min and max for each column, either from dataset (more accurately we should use the original range of full training data) or from documentation definitions if available there. 

Found some of these ranges by following link in _Dataset Description_ document (pdf) to https://developer.spotify.com/documentation/web-api/reference/#endpoint-get-audio-features and scroll/find to "AudioFeaturesObject" subheading for each Key field, data type and its range mostly are noted in descriptions.

For others I'm taking the min and max of the raw training dataset to get a range.
Did a bit more research on some variables like Tempo and Time Signature which generally correlate to one another.
For a Tempo range, the raw training dataset went as high as over the Prestissimo—extremely fast (more than 200 BPM) basic known top tempo marking, and down to 0, so I chose a range to start as 0 to 220 (Larghissimo—very, very slow, almost droning (20 BPM and below) is the lowest basic tempo marking published at my research source where you can read more: https://www.masterclass.com/articles/music-101-what-is-tempo-how-is-tempo-used-in-music#what-is-casual-musical-language). Note that our graphed dataset of least skipped may not extend well into either end of these ranges derived from the raw dataset, but are an accurate starting point and potential ranges for new values if refreshing analysis also.

Their Time Signature is listed as (meter) therefore it can be understood to represent the beats per measure, e.g. a duple time (2/2, 2/4, etc.) would be listed here as '2'. It's rare to see those below 2 or above 4 (quadruple) which are not an equivalent to one in that range, according to research source Liberty Park Music, but the max in dataset is 5 so we'll use 0 to 5 for our axis scale (source: https://www.libertyparkmusic.com/musical-time-signatures/#:~:text=There%20are%20three%20which%20are,has%20four%20beats%20per%20measure.)

```{r view raw track feautes dataset min}
apply(track_features,2,min)
```
```{r view raw track features dataset max}
apply(track_features,2,max)
```



```{r}
# Add 2 lines to least skipped radar dataframe: min and max of each column to show on plot.

# First because we have different ranges, we'll need to specify each in a dataframe of their own.
# Determined from aproximate min/max in training set: duration, release_year, us_popularity_estimate, key, loudness (backed up by research), mechanism, organism (those 2 were not listed in Spotify docs I found at all), energy, flatness, time_signature
# Note acoustiness, energy max, instrumentalness in dataset were 9.9... but docs indicate each is a 0.0 to 1.0 scale aligned with other columns; I'll test both for chart effectiveness (likely a data anomoly/error(s) exist outside above 1.0).
# Switched the erroneous 0 to 10 ranges to the 0 to 1 listed in docs because test proved that none in our dataset appeared away from the center of radar chart at 0. Once I disabled scientific notation, instrumentalness dropped values down below 1.0 in our dataset.
# Dynamic range also is accounted for as part of energy calculation so if its range is ineffective visually with the others, we could omit that column as its own variable as somewhat redundant.
# Updated to remove: dyn_range_mean = c(35.0, 0.0).
# Loudness in max aggregation shows -60 as the max, but as a natural scale I flipped min to that negative end of range whereas 0 became the max (and 0 was min in R aggregate function calculation).
# However, after testing, I switched loudness back to max as -60 because this allowed the center of radar range point to be 0 for all except the year, which makes for a more congruent view across variables with different range end points.
max_min <- data.frame(
  duration = c(2000.0, 0.0), release_year = c(2021, 1950), us_popularity_estimate = c(100.0, 0.0), acousticness = c(1.0, 0.0), beat_strength = c(1.0, 0.0), bounciness =  c(1.0, 0.0), danceability = c(1.0, 0.0), energy = c(1.0, 0.0), flatness = c(1.0, 0.0), instrumentalness = c(1.0, 0.0), key = c(11, 0), liveness = c(1.0, 0.0), loudness = c(-60.0, 0.0), mechanism = c(1.0, 0.0), organism = c(1.0, 0.0), speechiness = c(1.0, 0.0), tempo = c(220.0, 0.0), time_signature = c(5, 0), valence = c(1.0, 0.0)
)
rownames(max_min) <- c("Max", "Min")
# Bind the variable ranges to the data.
tracks_least_skipped_radar <- rbind(max_min, tracks_least_skipped_radar)
# Display the dataframe result.
tracks_least_skipped_radar
```
## Compare Least vs. Most Skipped Tracks as Radar Charts

_Ordered sequentially as steps taken;_ go straight to *Show Me the Side-by-Side Comparison* button to jump to the most effective visualization which will be incorporated in the data product to convey recommendation to our musician."

We'll first plot the default chart format _for illustrative purposes only_, then add some customization.

_**Produces a good start, but...**_ not quite what will help us undertand its contents easily in this case.
The default colors are a quantitative selection, which in our cases isn't really as necessary since we already have isolated similarity among our selected tracks based on them being the least often skipped.
Spotting any overall trends among features is more what we're going for here, and we'll adjust color to simplify this view more to support that goal.

```{r tracks_least_skipped_radar-default-plot-ex, dev='png', fig.show='hide'}
# FOR ILLUSTRATIVE PURPOSES ONLY! Lots of visual issues with this first try!
radarchart(tracks_least_skipped_radar)
```
```{r add-rownames}
# Add rownames back on for interactivity or if we split this into separate radar charts for each later.
# This error seems to be because the percents are floats not integers: Error in `.rowNamesDF<-`(x, value = value) : invalid 'row.names' length
#rownames(tracks_least_skipped_radar) <- paste(tracks_least_skipped_features[,"not_skipped_pct"], "% Skipped")
```

**Managing Shapes for Audience Perception**

We want to limit overinflated understanding of audio features based on the shape sizes that are perceived larger when adjacent categories differ considerably in value. In a chart like this with circular axes, the shape created grows in multiple dimensions related to variable on each of its sides as well as its placement on each axis/variable. This is different from e.g. a bar chart which is linear and will only increase in size in one direction along its axis. The risk with a radar chart is that we perceive greater emphasis on a data variable if it takes up more space of the polygon, i.e. it appears larger at that column data's axis than perhaps is meaningful. 

Another way to look at that is two axis with the same plotted point, may not look the same size or importance to the audience if one of them is next to points on their axes at a similar place, which would result in a 'smaller' shape there, compared to if one of them is next to points at very different places on their axes which requires a longer line or distance between points to close the shape and therefore could look 'larger' to the audience carrying greater perceptive emphasis with their understanding and conclusions from the graph.

Since the polygon shape can impact how we perceive sameness and different in magnitude on a radar graph, let's reorder the columns to put same and similar scales in neighborhoods, start at the top with most emphasis--I'll use popularity estimate, which look to all be above 90%. After some preliminary testing from viewing the default radar chart in our initial columns order, we are able to create and control a more visually accurate chart within this design, practicing our intention to limit misperceptions or our own misdirections in further analysis.

```{r adjust-columns-order}
# Note fmsb package plots radar in counter clockwise order, so we'll order our vector of column names according to what should appear last (12 o'clock) to 9:00 to 6:00 to 3:00, etc.
# Updated to remove "dyn_range_mean".
col_order <- c("us_popularity_estimate", "release_year", "duration", "time_signature", "tempo", "key", "loudness", "acousticness", "beat_strength", "bounciness", "danceability", "liveness", "mechanism", "organism", "energy", "flatness", "instrumentalness", "speechiness", "valence")

tracks_least_skipped_radar2 <- tracks_least_skipped_radar[, col_order]
# Show the datatable
tracks_least_skipped_radar2
```

```{r tracks_least_skipped_radar-plot, dev='png', fig.show='hide'}
# set a color palette variable for some greens that will be understood as a relative group (of the least skipped), and remain somewhat visible as distinct tracks individually also.
palette_greens <- brewer.pal(8,"Greens")
# for the fill of polygons, let's add some transparency with alpha
palette_greens_in <- alpha(palette_greens,0.2)

# Customize the radar chart
radarchart(tracks_least_skipped_radar2, axistype = 3,
           #add a title
           title = "Features of Least Skipped Spotify Tracks",
           #custom polygon, let's generalize all of these 'good' i.e. least skipped as a similar range of color using a R palette, since we'll compare it to 'bad' most skipped, not among themselves as much
           pcol = palette_greens, #colors the lines
           pfcol = palette_greens_in, #adds same colors as fill to polygons
           
           #custom the grid
           cglcol = "grey", cglty = 1, axislabcol = "grey30", cglwd = 0.8,
           
           #custom labels
           vlabels=c("US Popularity est. %", "Release Year", "Duration (ms)", "Time Signature", "Tempo", "Key", "Loudness", "Acousticness", "Beat Strength", "Bounciness", "Danceability", "Liveness", "Mechanism", "Organism", "Energy", "Flatness", "Instrumentalness", "Speechiness", "Valence"),
           vlcex = 0.725,
           #plot just a zero in center because all scales share min as 0
           caxislabels=c("0", "", "", "", "")
           )
```

**Mirror Preparation of Most Skipped Musical Tracks**
```{r take head of training sample}
tracks_most_skipped_features <- head(track_plays_skips_features)
tracks_most_skipped_features
```

Now, I'll repeat all that data prep and column order I did with the least skipped tracks to achieve a comparable result for a new radar chart.
_Note this also would be a great scenario to create one or more functions for easier repeatability for consistent charting in the future._
```{r prepare most skipped for radarchart}
# Remove unnecessary/non-quantitative columns
drop <- c("not_skipped_pct","track_id_clean","plays","skip_1","skip_2","skip_3","not_skipped","dyn_range_mean","mode","acoustic_vector_0","acoustic_vector_1","acoustic_vector_2","acoustic_vector_3","acoustic_vector_4","acoustic_vector_5","acoustic_vector_6","acoustic_vector_7")
tracks_most_skipped_radar = tracks_most_skipped_features[,!(names(tracks_most_skipped_features) %in% drop)]

# add the max and min to data frame, same value ranges as in the least skipped tracks
max_min <- data.frame(
  duration = c(2000.0, 0.0), release_year = c(2021, 1950), us_popularity_estimate = c(100.0, 0.0), acousticness = c(1.0, 0.0), beat_strength = c(1.0, 0.0), bounciness =  c(1.0, 0.0), danceability = c(1.0, 0.0), energy = c(1.0, 0.0), flatness = c(1.0, 0.0), instrumentalness = c(1.0, 0.0), key = c(11, 0), liveness = c(1.0, 0.0), loudness = c(-60.0, 0.0), mechanism = c(1.0, 0.0), organism = c(1.0, 0.0), speechiness = c(1.0, 0.0), tempo = c(220.0, 0.0), time_signature = c(5, 0), valence = c(1.0, 0.0)
)
rownames(max_min) <- c("Max", "Min")
# Bind the variable ranges to the data.
tracks_most_skipped_radar <- rbind(max_min, tracks_most_skipped_radar)
# Display the dataframe result.
#tracks_most_skipped_radar

# Reorder the columns for radar sequence.
col_order <- c("us_popularity_estimate", "release_year", "duration", "time_signature", "tempo", "key", "loudness", "acousticness", "beat_strength", "bounciness", "danceability", "liveness", "mechanism", "organism", "energy", "flatness", "instrumentalness", "speechiness", "valence")
tracks_most_skipped_radar2 <- tracks_most_skipped_radar[, col_order]
# Show the datatable
tracks_most_skipped_radar2
```
```{r tracks_most_skipped_radar2-plot, dev='png', fig.show='hide'}
# set a color palette variable for some greys this time that will be more distant and still understood as a relative group (of the least skipped), and remain somewhat visible as distinct tracks individually also, like the least skipped.
palette_greys <- brewer.pal(8,"Greys")
# for the fill of polygons, let's add some transparency with alpha
palette_greys_in <- alpha(palette_greys,0.2)

# Customize the radar chart
radarchart(tracks_most_skipped_radar2, axistype = 3,
           #add a title
           title = "Features of Most Skipped Spotify Tracks",
           #custom polygon, let's generalize all of these 'bad' i.e. most skipped as a similar range of color using a R palette, less emphasis than the greens were for 'good' tracks, since we'll compare it to 'good' least skipped, not among themselves as much.
           pcol = palette_greys, #colors the lines
           pfcol = palette_greys_in, #adds same colors as fill to polygons
           
           #custom the grid; we will test greys on grey grid and adjust if needed.
           cglcol = "grey", cglty = 1, axislabcol = "grey30", cglwd = 0.8,
           
           #custom labels
           vlabels=c("US Popularity est. %", "Release Year", "Duration (ms)", "Time Signature", "Tempo", "Key", "Loudness", "Acousticness", "Beat Strength", "Bounciness", "Danceability", "Liveness", "Mechanism", "Organism", "Energy", "Flatness", "Instrumentalness", "Speechiness", "Valence"),
           vlcex = 0.725,
           #plot just a zero in center because all scales share min as 0
           caxislabels=c("0", "", "", "", "")
           )
```
## Show Me the Side-by-Side Comparison!

Now we will plot the least and most skipped charts beside one another for comparison.
```{r tracks_least_most_skipped_radar2-plots, dev='png', fig.show='animate'}
opar <- par() 
# Define settings for plotting in a 1x2 (2 column) grid, with appropriate margins:
par(mar = rep(.8,4)) #may require testing to get margins between plots looking good
par(mfrow = c(1,2)) # 1 row, 2 columns
# Produce a radar-chart for each desired plot again, least skipped at left; most to right
# Customized least skipped radar chart
radarchart(tracks_least_skipped_radar2, axistype = 3,
           #add a title
           title = "Features of Least Skipped Spotify Tracks",
           #custom polygon, let's generalize all of these 'good' i.e. least skipped as a similar range of color using a R palette, since we'll compare it to 'bad' most skipped, not among themselves as much
           pcol = palette_greens, #colors the lines
           pfcol = palette_greens_in, #adds same colors as fill to polygons
           
           #custom the grid
           cglcol = "grey", cglty = 1, axislabcol = "grey30", cglwd = 0.8,
           
           #custom labels
           vlabels=c("US Popularity est. %", "Release Year", "Duration (ms)", "Time Signature", "Tempo", "Key", "Loudness", "Acousticness", "Beat Strength", "Bounciness", "Danceability", "Liveness", "Mechanism", "Organism", "Energy", "Flatness", "Instrumentalness", "Speechiness", "Valence"),
           vlcex = 0.635,
           #plot just a zero in center because all scales share min as 0
           caxislabels=c("0", "", "", "", "")
           )
# Customized most skipped radar chart
radarchart(tracks_most_skipped_radar2, axistype = 3,
           #add a title
           title = "Features of Most Skipped Spotify Tracks",
           #custom polygon, let's generalize all of these 'bad' i.e. most skipped as a similar range of color using a R palette, less emphasis than the greens were for 'good' tracks, since we'll compare it to 'good' least skipped, not among themselves as much.
           pcol = palette_greys, #colors the lines
           pfcol = palette_greys_in, #adds same colors as fill to polygons
           
           #custom the grid; we will test greys on grey grid and adjust if needed.
           cglcol = "grey", cglty = 1, axislabcol = "grey30", cglwd = 0.8,
           
           #custom labels
           vlabels=c("US Popularity est. %", "Release Year", "Duration (ms)", "Time Signature", "Tempo", "Key", "Loudness", "Acousticness", "Beat Strength", "Bounciness", "Danceability", "Liveness", "Mechanism", "Organism", "Energy", "Flatness", "Instrumentalness", "Speechiness", "Valence"),
           vlcex = 0.635,
           #plot just a zero in center because all scales share min as 0
           caxislabels=c("0", "", "", "", "")
           )
# Restore the standard par() settings
par <- par(opar) 
```
## Conclusions of Exploratory Data Analysis
Spotify users show trends of preferences toward playing songs through entirety with:

1. Faster Tempo, Higher Energy songs
2. Higher Beat Strengths, Bounciness and Danceability
3. Low Acousticness with sounds which are more Mechanical than Organic in nature

_Limitations:_ The most skipped (head by % not skipped of played) and bottom (tail of same) are illustrative for comparing patterns in differences of song features, but may not indicate statistically significant findings.

A categorical or neighborhood model could further assess these findings with a confidence level.

### Data & Cited Sources

Sample data contains over 300 unique tracks each played 50 times or more as analyzed in this notebook, of the _Training_Set_And_Track_Features_Mini (17.2 MB)_ from AI Crowd Spotify Sequential Skip Prediction Challenge By Spotify, which sponsored the AI Crowd challenge (https://www.spotify.com/).

**Data source(s):** https://aicrowd-production.s3.eu-central-1.amazonaws.com/dataset_files/challenge_204/16772e7f-7871-4d42-a44f-5f399f40fd94_training_set_track_features_mini.tar.gz?X-Amz-Algorithm=AWS4-HMAC-SHA256&X-Amz-Credential=AKIAJ6IZH6GWKDCCDFAQ%2F20210415%2Feu-central-1%2Fs3%2Faws4_request&X-Amz-Date=20210415T034213Z&X-Amz-Expires=86400&X-Amz-SignedHeaders=host&X-Amz-Signature=8f355b250186ffef249ff042a138717684a8aace6732c6804ade55e22ce21a7d

**Image source(s):** 
Spotify Icon: <a href="https://iconscout.com/icons/spotify" target="_blank">Spotify Icon</a> by <a href="https://iconscout.com/contributors/Hopstarter" target="_blank">Jojo Mendoza</a>

Next Icon: <a href="https://iconscout.com/icons/next" target="_blank">Next Icon</a> by <a href="https://iconscout.com/contributors/saepul-nahwan">Saepul Nahwan</a> on <a href="https://iconscout.com">Iconscout</a>

Fast Icon: <a href="https://iconscout.com/icons/fast" target="_blank">Fast Icon</a> by <a href="https://iconscout.com/contributors/taufik">Taufik Ramadhan</a> on <a href="https://iconscout.com">Iconscout</a>

Headphone Icon: <a href="https://iconscout.com/icons/headphone" target="_blank">Headphone Icon</a> by <a href="https://iconscout.com/contributors/fatkhulkarim" target="_blank">Fatkhul Karim</a>

**Cited: **
All last accessed with published notebook date except where otherwise noted.

Produced with instructive direction and initial coding starts from Justin Elszasz "<a href=mailto:justin.elszasz@baltimorecity.gov>justin.elszasz@baltimorecity.gov</a> as part of MICA Data Analytics & Visualization program course Visual Storytelling.

Radar Chart guidelines/bestpractices and FSMB package requirements: https://www.r-graph-gallery.com/spider-or-radar-chart.html

The Radar Chart and its Caveats, A collection of common dataviz caveats by Data-to-Viz.com: https://www.data-to-viz.com/caveat/spider.html

Spotify Developer documentation: https://developer.spotify.com/documentation/web-api/reference/#endpoint-search
See especially terms under section "AudioFeaturesObject".

Reserach info on Time Signature: https://www.libertyparkmusic.com/musical-time-signatures/#

 and Tempo: https://www.masterclass.com/articles/music-101-what-is-tempo-how-is-tempo-used-in-music#what-is-casual-musical-language

**R Session Info**
```{r session_info}
sessionInfo()
```

